---
title: "Ipomopsis temperature experiments: Floral traits"
author: "Carrie Wu, John Powers, David Hopp, Diane Campbell"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE 
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
.main-container { max-width: 1000px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(lme4)
library(lmerTest)
library(emmeans)
library(broom)
library(viridis)
library(ggsignif)
library(RColorBrewer)
knitr::opts_chunk$set(comment="", cache=T, warning=F, message=F, 
                      fig.path = "plots-traits/", dev="svglite", dev.args=list(fix_text_size=FALSE), 
                      fig.height=5, fig.width=5)
```

```{r read_data}
# Load in all the Ipomopsis temperature experiment data
# Includes census, phenology, floral and leaf traits, and seeds
# census -------------------------------------------------------------------

census.maxf <- read_csv("data/2021 Maxfield Rosettes - 2021OTCs.csv") %>% 
  mutate(across(c(plant, plot), as.character),
         temp=fct_relevel(fct_recode(temp,"warmed"="OTC"),"control"))

#TODO look at survival and flowering from 2021 to 2022 (2021 Maxfield Rosettes - 2021.csv)

# floral traits --------------------------------------------------------------

morph <- bind_rows(GNA=read_csv("data/traits/GNA floral traits - 2021.csv", na="."),
                   SC=read_csv("data/traits/Spring Creek floral traits - 2021.csv", na="."), .id="site") %>% 
  rename(plantid = plant, temp = location, plotid=OTC) %>% drop_na(temp) %>% 
  mutate(temp=fct_relevel(fct_recode(temp,"warmed"="in", "control"="out"),"control")) %>% 
  bind_rows(read_csv("data/traits/2021 Maxfield Floral Traits - Morphology21.csv") %>% 
              mutate(plant = as.character(plant), site="Maxfield") %>% 
              left_join(census.maxf) %>% drop_na(plot)) %>% 
  mutate(species = if_else(site=="SC", "ten", "agg"), year=year(date))

nectar <- read_csv("data/traits/2021 Maxfield Floral Traits - Nectar21.csv") %>% 
  mutate(across(c(plant), as.character)) %>% 
  left_join(census.maxf) %>% 
  mutate(nectar_volume=nectar_48_h_mm*5/32,
         nectar_sugar=nectar_volume*nectar_conc,
         plantid=paste0(plotid,plant), 
         site="Maxfield", species="agg", year=year(date))

flrcol.raw <- read_csv("data/traits/2021 Maxfield Flower Color - 2021.csv", name_repair="minimal") %>% as.data.frame()
wl <- flrcol.raw[,1] #350:1000
flrcol.raw <- flrcol.raw[,seq(2,length(flrcol.raw),2)]

flrcol.long <- cbind(wl,stack(flrcol.raw)) %>% 
  mutate(red=wl>=626&wl<=700, #sum red and green regions to calculate redness index
         green=wl>=476&wl<=550,
         total=wl>=401&wl<=750,
         color=ifelse(red,"red",ifelse(green,"green",NA)))
flrcol.total <- flrcol.long %>% group_by(color,total,ind) %>% summarize(values=sum(values))

flrcol <- pivot_wider(flrcol.total,names_from=c("color","total"),values_from="values") %>% 
  mutate(total = green_TRUE+red_TRUE+NA_TRUE,
         redness = (red_TRUE-green_TRUE)/total,
         ind = str_replace(ind,fixed("."),"_")) %>% 
  separate(ind, into=c("plantid","rep"), sep="_", fill="right", remove=F) %>%
  mutate(rep=as.integer(replace_na(rep, "0"))+1) %>% 
  left_join(census.maxf) %>% mutate(site="Maxfield", species="agg")

# phenology ---------------------------------------------------------------

phen.raw <- read_csv("data/traits/2021 Maxfield Phenology - 2021.csv")%>% 
  mutate(plot = as.character(plot), 
         plotid = paste0(plot, subplot),
         plantid = paste0(plotid, plant),
         plant = as.character(plant),
         julian = yday(date),
         year="2021",
         height_cm=as.integer(as.character(height_cm)),
         open = rowSums(select(., starts_with("open")), na.rm=T),
         buds = rowSums(select(., starts_with("buds")), na.rm=T),
         eggs = rowSums(select(., starts_with("eggs")), na.rm=T)) %>% 
  left_join(census.maxf) %>% 
  mutate_if(is.character, as.factor) 

#some plants were measured late, merge with the earlier date
phen <- phen.raw %>% mutate(date=ymd(recode(as.character(date),
                                            "2021-07-02"="2021-06-30", "2021-07-26"="2021-07-20")),
                            julian=yday(date)) %>% 
  complete(nesting(plantid, plotid, plot, subplot, snow, temp),nesting(julian, date), fill=list(open=0,buds=0)) %>% #add zeros to weeks the plant was not counted
  mutate(flowering = open + buds > 0,
         has_egg = eggs > 0, 
         site="Maxfield", species="agg", year=year(date))

# seeds -------------------------------------------------------------------

counts_19_20 <- c("seeds","fruits","fruits_split","aborts","fruits_fly_no_seeds","fruits_fly_with_seeds","seeds_fly","fruits_caterpillar","fruits_early_uncountable","flowers_buds","flowers_buds_collected_early","flowers_buds_collected_last")
sds21.date <- read_csv("data/traits/2021 Maxfield Seeds - 2021.csv", skip=1) %>% 
  mutate(plot = as.character(plot), 
         plotid = paste0(plot, subplot),
         plantid = paste0(plotid, plant),
         plant = as.character(plant),
         julian = factor(yday(date)),
         year = year(date), plantid = paste0(plotid,plant),
         species="agg", site="Maxfield") %>% 
  mutate(flowers_buds_collected_last = ifelse(yday(date) >= 235, flowers_buds, 0), #8/23
         flowers_buds_collected_early = ifelse(yday(date) < 235, flowers_buds, 0)) %>% 
  mutate(across(all_of(counts_19_20), replace_na, 0)) %>% 
  left_join(census.maxf) %>% drop_na(temp) %>% #8 plants with no census record (also not in traits)
  mutate_if(is.character, as.factor) 

sds21 <- sds21.date %>% 
  group_by(across(c("plant", "plantid", colnames(census.maxf)))) %>% 
  summarize(across(all_of(counts_19_20), sum, na.rm=T), dates=n(), .groups="drop")

sds21.date.zeroed <- sds21.date %>% 
  filter(date %in% (sds21.date %>% drop_na(date) %>% count(date) %>% filter(n>2) %>% pull(date))) %>% 
  complete(nesting(plantid, plotid, plot, subplot, snow, temp), nesting(julian, date), fill=list(seeds=0, fruits=0, fruits_split=0, fruits_fly_no_seeds=0, fruits_fly_with_seeds=0, seeds_fly=0, fruits_caterpillar=0)) %>% 
  mutate(julian=as.integer(as.character(julian)),
         seeds_per_fruit = seeds/fruits,
         fruits_with_seeds = fruits + fruits_split + fruits_fly_with_seeds,
         fruits_nonaborted = fruits_with_seeds + fruits_fly_no_seeds + fruits_caterpillar + fruits_split,
         prop_infested	= (fruits_fly_no_seeds + fruits_fly_with_seeds + fruits_caterpillar) / fruits_nonaborted,
         fruiting = fruits_nonaborted > 0)

sds <- sds21 %>% 
  mutate(#flowers_volatiles = replace_na(flowers_volatiles, 0), #TODO get volatiles for 2018 & 2019
    #flowers_destroyed = flowers_morphology_nectar + flowers_volatiles, #TODO + flowers_color
    seeds_per_fruit	= seeds / fruits,
    fruits_aborted	= aborts + flowers_buds_collected_last,
    seeds_est	= seeds + seeds_fly + (flowers_buds_collected_early + #flowers_destroyed + 
                                       fruits_early_uncountable) * (seeds/(fruits + fruits_aborted + fruits_fly_with_seeds + fruits_fly_no_seeds + fruits_caterpillar)) + fruits_split * seeds_per_fruit, 
    fruits_with_seeds	= fruits + fruits_split + fruits_fly_with_seeds, 
    fruits_nonaborted	= fruits_with_seeds + fruits_fly_no_seeds + fruits_caterpillar + fruits_split,
    flowers_est	= fruits_nonaborted + aborts + flowers_buds, #+ flowers_destroyed,# aborts used to be fruits_aborted, which includes flowers_buds_collected_last, but these are already in flowers_buds
    prop_infested	= (fruits_fly_no_seeds + fruits_fly_with_seeds + fruits_caterpillar) / fruits_nonaborted,
    prop_aborted	= fruits_aborted / (fruits_nonaborted + fruits_aborted),
    prop_nonaborted	= fruits_nonaborted / (fruits_nonaborted + fruits_aborted),
    seeds_per_flower = seeds_est / flowers_est) %>% 
  filter(flowers_est > 0)#one plant "gone"

write.csv(sds, "data/2021_Maxfield_sds.csv")

# soilmoisture ---------------------------------------------------------------

sm <- bind_rows(read_csv("data/2019 & 2021 SC & GNA Soil Moisture - general.csv"),
                read_csv("data/2021 Maxfield Soil Moisture - 2021.csv") %>% 
                  mutate(temp = ifelse(location %in% c("OTCV","OTCO") | !is.na(OTC),"warmed", "control"),
                         site="Maxfield"),
                read_csv("data/2021 Maxfield Leaf Traits - 2021round1.csv") %>% 
                  select(date=date_collected, plot, subplot, plant, temp, VWC, period=per_us) %>% 
                  mutate(temp = ifelse(temp=="OTC", "warmed", "control"), site="Maxfield")) %>% 
                  mutate(year=year(date))

# meltdates ---------------------------------------------------------------

# 2021 Maxfield HOBO threshold=10,000 lux
# normal: plots
# 1 = 21-04-30 
# 3 = 21-04-30
# 6 = 21-05-02
# early plots
# 2 = 21-04-14
# 4 = 21-04-14
# 5 = 21-04-12

meltdates <- tibble::tribble( ~year, ~sun_date, # average light level melt day in normal snowmelt plots (1,3,6) at Maxfield
                              2018,      119, # from Maxfield Results - timings sheet
                              2019,      158, # threshold = 10,000 lux
                              2021,      121) %>% # Manually from lab drive: 2021 Maxfield Hobo overwinter
  mutate(site="Maxfield")

# names --------------------------------------------------------------

spnames <- c(agg="I. aggregata", ten="I. tenuituba")
sitenames <- c(GNA="Gothic Natural Area",Maxfield="Maxfield Meadow",SC="Spring Creek")
year_pal <- setNames(brewer.pal(3, name="Set2")[c(2,3,1)], meltdates$year)

traits <- c("corolla_length", "corolla_width", 
            "nectar_volume", "nectar_conc",
            "height_cm","redness",
            "open","seeds_per_fruit", "seeds_est", "fruits_nonaborted",  "flowers_est", 
            "prop_infested", "prop_aborted", "seeds_per_flower") 
floraltraits <- traits[1:6]
seedtraits <- traits[7:14] 
traitnames.units <- set_names(c(
  "Corolla length (mm)", "Corolla width (mm)",
  "Nectar production (\U00B5L/day)", "Nectar conc. (% by mass)",
  "Inflorescence height (cm)", "Redness index", 
  "Open flowers", "Seeds per fruit", "Estimated total seeds", "Nonaborted fruits", "Flower number", 
  "Prop. nonaborted fruits infested","Prop. fruits aborted","Estimated seeds per flower"), traits)

traitnames <- set_names(c(
  "Corolla length", "Corolla width",
  "Nectar production", "Nectar conc.",
  "Inflorescence height", "Redness index", 
  "Open flowers", "Seeds per fruit", "Estimated total seeds", "Nonaborted fruits", "Flower number", 
  "Prop. nonaborted fruits infested","Prop. fruits aborted","Estimated seeds per flower"), traits)
```

# Sample sizes

## Timings

```{r timings, fig.height=6}
timings <- bind_rows(
  meltdates=meltdates %>% rename(begin=sun_date) %>% mutate(end=begin),
  otc= read_csv("./data/CWu OTC Hobos - duration.csv") %>% 
    mutate(begin=yday(OTC_start)-1, end=yday(OTC_end)+1, .keep="unused"), #start/end were trimmed 1 day 
  sm = sm %>% group_by(site, year) %>% summarize(begin = min(yday(date)), end = max(yday(date))),
  vt = read_csv("data/volatiles/CWu_meta21.csv") %>% rename(date=Date, site=Site) %>% mutate(year=year(date)) %>%
    filter(Tissue=="flower", !is.na(site)) %>% group_by(site, year) %>% 
    summarize(begin = min(yday(date)), end = max(yday(date))),
  mt = morph %>% drop_na(corolla_length) %>% group_by(site, year) %>% 
    summarize(begin = min(yday(date), na.rm=T), end = max(yday(date), na.rm=T)),
  nt = nectar %>% group_by(site, year) %>% summarize(begin = min(yday(date), na.rm=T), end = max(yday(date), na.rm=T)),
  flrcol = data.frame(year=2021, site="Maxfield", begin=yday(ymd("2021-07-14")), end=yday(ymd("2021-07-26"))),
  ph = phen %>% drop_na(height_cm) %>% group_by(site, year) %>% summarize(begin = min(yday(date)), end = max(yday(date))),
  sds = sds21.date %>% group_by(site, year) %>% summarize(begin = min(yday(date), na.rm=T), end = max(yday(date), na.rm=T)), 
  .id="variable") %>% ungroup

timings_labels <- c(meltdates="Snowmelt", otc="Open top chambers",
                   vt="Volatiles",
                   sm="Soil moisture",         ph="Fly eggs", #Inflorescence height
                   mt="Floral morphology",     nt="Nectar", 
                   sds="Seeds", flrcol="Flower color")

timings %>% mutate(range=paste(begin,end,sep=" - "), .keep="unused") %>% 
  pivot_wider(names_from=year, values_from=range, names_sort=T)

dodging <- position_dodge2(width=1, reverse = T, preserve="single")
timings %>% filter(site!="bb") %>% 
  mutate(variable=fct_reorder(variable, begin, na.rm=T, .desc=T), year=factor(year)) %>% 
  ggplot(aes(y=variable, color=year, shape=site))+ 
  geom_linerange(aes(xmin=begin, xmax=end), 
                 position = dodging, size=1.2, show.legend=FALSE) + 
  geom_point(aes(x=begin), position = dodging, size=2.5)+
  geom_point(aes(x=end), position = dodging, size=2.5) + 
  scale_y_discrete("", labels=timings_labels)+
  scale_x_continuous("Day of year", breaks=seq(80,240, by=20)) +
  labs(color="Year", shape="Site")+ scale_color_manual(values=year_pal)+
  theme_minimal() + theme(legend.position = "top", text=element_text(size=14, color="black"), 
                          axis.text = element_text(color="black"),
                          panel.border = element_rect(fill=NA, colour = "black"), legend.box="vertical")
```


## Measurements per plant
```{r n_per_plant}
morph %>% drop_na(corolla_length) %>% count(site, plantid, name="n_per_plant") %>% count(site, n_per_plant) %>%
  kable(caption="corolla_length")
nectar %>% drop_na(nectar_volume) %>% count(site, plantid, name="n_per_plant") %>% count(site, n_per_plant) %>% 
  kable(caption="nectar_volume")
phen %>% drop_na(height_cm) %>% count(site, plantid, name="n_per_plant") %>% count(site, n_per_plant) %>% 
  kable(caption="height_cm")
flrcol %>% drop_na(redness) %>% count(site, plantid, name="n_per_plant") %>% count(site, n_per_plant) %>% 
  kable(caption="redness")
```

## Number of plants
```{r n_total}
morph %>% drop_na(corolla_length) %>% count(site, temp, plantid) %>% count(site, temp) %>% kable(caption="corolla_length")
nectar%>% drop_na(nectar_volume)  %>% count(site, temp, plantid) %>% count(site, temp) %>% kable(caption="nectar_volume")
phen  %>% drop_na(height_cm, temp)  %>% count(site, temp, plantid) %>% count(site, temp) %>% kable(caption="height_cm")
flrcol%>% drop_na(redness, temp)    %>% count(site, temp, plantid) %>% count(site, temp) %>% kable(caption="redness")

#Average data by plant
morph.mean <- morph %>% drop_na(corolla_length) %>% 
  group_by(species, site, temp, snow, plotid, plantid) %>%
  summarize(across(c(corolla_length, corolla_width), mean, na.rm=T), n = n())
nectar.mean <- nectar %>% drop_na(nectar_volume) %>% 
  group_by(species, site, temp, snow, plotid, plantid) %>%
  summarize(across(c(nectar_volume, nectar_conc, nectar_sugar), mean, na.rm=T), n = n())
phen.max <- phen %>% drop_na(height_cm, temp) %>% 
  group_by(species, site, temp, snow, plotid, plantid) %>% 
  summarize(height_cm=max(height_cm, na.rm=T)) #use the maximum plant height instead of the mean across the year (deer browsing)
flrcol.mean <- flrcol %>% drop_na(redness, temp) %>% 
  group_by(species, site, temp, snow, plotid, plantid) %>% 
  summarize(redness=mean(redness, na.rm=T))

all.mean <- morph.mean %>% rename(n.morph=n) %>% 
  full_join(rename(nectar.mean, n.nectar=n)) %>% full_join(phen.max) %>% full_join(flrcol.mean) %>% full_join(sds)
traits.sd <- map_dbl(traits,     ~sd(all.mean[[.x]], na.rm=T)) %>% set_names(traits)
```

# Soil moisture

```{r soilmoisture, fig.width=7}
ggplot(sm %>% filter(date!=ymd("2021-06-23")), #before Maxfield OTCs started
       aes(x=paste(year(date), site,format(date, "%m-%d")), y=VWC, color=temp)) +
  geom_boxplot(outlier.size = 0.5)+
  theme_minimal() + theme(axis.text.x=element_text(angle=90, hjust=0), legend.position="top", 
                          axis.title.x = element_blank()) + 
  scale_color_brewer(palette = "Set1", direction=-1)+
  labs(y="Soil Volumetric Water Content (%)", x="", color="Temperature")
```

# Trait correlations
```{r corr}
ggplot(morph, aes(shape=site, y=corolla_width, x=corolla_length, color=temp)) + 
  facet_wrap(vars(species), labeller = as_labeller(spnames)) +
  geom_point() + geom_smooth(aes(group=species), color="black") + 
  labs(x="Corolla length", y="Corolla width", shape="Site", color="Temperature") + 
  scale_color_brewer(palette="Set1", direction=-1) + theme_bw() + 
  theme(strip.text=element_text(face="italic"))
ggplot(nectar, aes(y=nectar_conc, x=nectar_volume, color=temp)) + 
  geom_point() + geom_smooth(color="black") + 
  labs(x="Nectar volume (uL)", y="Nectar concentration", shape="Site", color="Temperature") + 
  scale_color_brewer(palette="Set1", direction=-1) + theme_bw() + 
  theme(strip.text=element_text(face="italic"))
```

# Trait plasticity
```{r plasticity_multisite, fig.height=7.5}
morph.mean %>% pivot_longer(c(corolla_length, corolla_width)) %>% 
ggplot(aes(x=temp, y=value, color=temp)) + #paste(spnames[species], sitenames[site], sep="\n")
  facet_grid(name~site, scales="free_y", labeller = as_labeller(c(traitnames.units, sitenames)))+
  geom_violin(draw_quantiles=c(0.25, 0.5, 0.75)) + 
  geom_signif(comparisons = list(c(1,2)), test="t.test",
              map_signif_level=TRUE, color="black",size=0.4, textsize=3) +
  geom_point(position=position_jitterdodge(dodge.width=0.9, jitter.width=0.1), size=1) +
  labs(x="", y="", color="Temperature") + 
  scale_y_continuous(limits=c(0,NA))+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal() +
  theme(legend.position = "top", axis.title.x=element_blank(), axis.text.x=element_blank(),
        panel.grid.major.x = element_blank())
```

```{r plasticity, fig.width=7.5}
all.mean %>% pivot_longer(any_of(floraltraits[-c(1:2)])) %>% 
  filter(site=="Maxfield") %>% 
  drop_na(value) %>% 
ggplot(aes(x=temp, y=value, color=temp)) + 
  facet_wrap(vars(name), scales="free_y", labeller = as_labeller(traitnames.units), nrow=1)+
  geom_violin(draw_quantiles=c(0.25, 0.5, 0.75)) + 
  geom_point(position=position_jitterdodge(dodge.width=0.9, jitter.width=0.1), size=1) +
  geom_signif(comparisons = list(c(1,2)), test="t.test",
              map_signif_level=TRUE, color="black",size=0.4, textsize=3) +
  labs(x="", y="", color="Temperature") + 
  scale_y_continuous(limits=c(0,NA))+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal() +
  theme(legend.position = "top", axis.title.x=element_blank(), axis.text.x=element_blank(),
        panel.grid.major.x = element_blank())

all.mean %>% pivot_longer(any_of(seedtraits)) %>% 
  filter(site=="Maxfield") %>% 
  drop_na(value) %>% 
ggplot(aes(x=temp, y=value, color=temp)) + 
  facet_wrap(vars(name), scales="free_y", labeller = as_labeller(traitnames.units), nrow=2)+
  geom_violin(draw_quantiles=c(0.25, 0.5, 0.75)) + 
  geom_point(position=position_jitterdodge(dodge.width=0.9, jitter.width=0.1), size=1) +
  geom_signif(comparisons = list(c(1,2)), test="t.test",
              map_signif_level=TRUE, color="black",size=0.4, textsize=3) +
  labs(x="", y="", color="Temperature") + 
  scale_y_continuous(limits=c(0,NA))+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal() +
  theme(legend.position = "top", axis.title.x=element_blank(), axis.text.x=element_blank(),
        panel.grid.major.x = element_blank())
```


# Effect of snowmelt

```{r snow}
lmer(corolla_length~temp*snow+(1|plotid)+(1|plantid),data=filter(morph, site=="Maxfield")) %>% anova %>%
  kable(caption="corolla_length")
lmer(corolla_width ~temp*snow+(1|plotid)+(1|plantid),data=filter(morph, site=="Maxfield")) %>% anova %>%
  kable(caption="corolla_width")
lmer(nectar_volume ~temp*snow+(1|plotid)+(1|plantid),data=nectar) %>% anova %>% 
  kable(caption="nectar_volume")
lmer(nectar_conc   ~temp*snow+(1|plotid)+(1|plantid),data=nectar) %>% anova %>% 
  kable(caption="nectar_conc")
lmer(height_cm   ~temp*snow+(1|plotid),data=phen.max) %>% anova %>% 
  kable(caption="height_cm")
lmer(redness     ~temp*snow+(1|plotid)+(1|plantid), data=flrcol) %>% anova %>% 
  kable(caption="redness")
```

# Effect of temperature

```{r temp}
lmer(corolla_length~temp*species+(1|plantid),data=morph) %>% anova %>% kable(caption="corolla_length")
lmer(corolla_width ~temp*species+(1|plantid),data=morph) %>% anova %>% kable(caption="corolla_width")
lmer(nectar_volume ~temp+(1|plantid),data=nectar)        %>% anova %>% kable(caption="nectar_volume")
lmer(nectar_conc   ~temp+(1|plantid),data=nectar)        %>% anova %>% kable(caption="nectar_conc")
lm(height_cm     ~temp, data=phen.max)      %>% anova %>% kable(caption="height_cm")
lmer(redness       ~temp+(1|plantid), data=flrcol)       %>% anova %>% kable(caption="redness")
```

# Flower color

## Spectra

```{r flrcol_uv}
library(pavo)
# use human range to estimate perceived color (requires interpolated 300 nm data)
flrspec.human <- cbind(wl, flrcol.raw) %>% as.rspec(lim = c(300,700))#no UV data below 350 nm, interpolate
#plot(flrspec.human, col = turbo(ncol(flrspec)), main="Raw data")
#plot(flrspec.human, col = spec2rgb(flrspec.human), main="Raw data")
flrspec.human.nm <- procspec(flrspec.human, opt = c("smooth", "max"), span=0.08) #smooth and normalize
#plot(flrspec.human.nm, col = turbo(ncol(flrspec)), main="Divide by max")
flrspec.human.nm.zero <- procspec(flrspec.human.nm, fixneg="zero") #throw out negative reflectance
#plot(flrspec.human.nm.zero, col = turbo(ncol(flrspec)), main="Divide by max")
#plot(flrspec.human.nm.zero, col = spec2rgb(flrspec.human.nm.zero), main="Divide by max")
flrcol$RGB <- spec2rgb(flrspec.human.nm.zero)
```

```{r flrcol, dev="png", dev.args=list()}
ggplot(flrcol.long, aes(x=wl, y=values, color=ind)) + geom_line() + 
  annotate("rect", xmin=401, xmax=750, ymin=0, ymax=100, fill=NA, color="black", size=1) +
  annotate("rect", xmin=476, xmax=550, ymin=0, ymax=100, fill=NA, color="green", size=1) +
  annotate("rect", xmin=626, xmax=700, ymin=0, ymax=100, fill=NA, color="red", size=1) +
  labs(x="Wavelength (nm)", y="Reflectance (%)")+ guides(color="none") + theme_minimal() +
  scale_x_continuous(n.breaks=10, limits=c(350, 1000)) + scale_color_viridis_d(option="turbo")

plot_colranges <- function() {
  abline(v=476, lwd=2, col="green"); abline(v=550, lwd=2, col="green")
  abline(v=626, lwd=2, col="red");   abline(v=700, lwd=2, col="red")
  abline(v=401, lwd=2, col="black"); abline(v=750, lwd=2, col="black")
}

flrspec <- cbind(wl, flrcol.raw) %>% as.rspec(lim = c(401,750))
plot(flrspec, col = turbo(ncol(flrspec)), main="Raw data"); plot_colranges()
heatmap(t(as.matrix(flrspec)[,-1]), Rowv = NA, Colv=NA, scale="none", col=viridis(128), main="Raw data") #block of samples zeroed wrong?
ggplot(flrcol.total %>% mutate(index=row_number()), aes(x=index, y=values, color=paste(color, total))) +
  geom_point() + geom_smooth(span=0.2) + labs(x="Sample number", y="Summed reflectance", color="Wavelength range")+
  scale_color_manual(values=c("green", "grey50" ,"black","red"), 
                     labels=c("green","outside 400-750 nm", "other inside 400-750 nm", "red"))
ggplot(flrcol %>% mutate(index=row_number()), aes(x=index, y=redness)) +
  geom_point() + geom_smooth(span=0.3) + labs(x="Sample number", y="Redness")

flrspec.nm.zero <- procspec(flrspec, opt = c("smooth", "max"), span=0.08, fixneg="zero") #smooth, normalize, zero negatives
plot(flrspec.nm.zero, col = turbo(ncol(flrspec)), main="Divide by max"); plot_colranges()
heatmap(t(as.matrix(flrspec.nm.zero)[,-1]), Rowv = NA, Colv=NA, scale="none", col=viridis(128), main="Divide by max") #block fixed by normalizing
plot(flrspec.nm.zero, col = flrcol$RGB, main="Divide by max"); plot_colranges()
plot(flrspec.nm.zero, col = viridis(100)[300*(flrcol$redness-0.1)], main="Divide by max, color by redness"); plot_colranges()
flrspec.sum.zero <- procspec(flrspec, opt = c("smooth", "sum"), span=0.08, fixneg="zero") #smooth, normalize, zero negatives
plot(flrspec.sum.zero, col = viridis(100)[300*(flrcol$redness-0.1)], main="Divide by sum, color by redness"); plot_colranges()
```

## PCA
```{r flrcol_pca}
flrspec.nm.zero.bin <- procspec(flrspec.nm.zero, opt = c("bin", "center"), bins=16) 
rownames(flrspec.nm.zero.bin) <- flrspec.nm.zero.bin[,1]
library(vegan)
flrcol.pca <- rda(t(flrspec.nm.zero.bin[,-1]), scale = TRUE)
flrcol <- bind_cols(flrcol, scores(flrcol.pca, display="sites"))
plot(flrcol.pca, type="none", main=paste("PCA of", nrow(flrspec.nm.zero.bin),"wavelength bins"))
ordispider(flrcol.pca, flrcol$plantid, col="grey40")
points(flrcol.pca, display="sites", col=flrcol$RGB, pch=c(17,19)[flrcol$temp])
text(flrcol.pca, display="species", col="black", cex=0.8)

library(GGally)
ggpairs(select(flrcol, total, red_TRUE, green_TRUE, redness, PC1), 
        lower=list(continuous=wrap("smooth", method = "loess")))
```

## CAP
```{r flrcol_cap}
flrcol.cap <- capscale(t(flrspec.nm.zero.bin[,-1]) ~ temp, scale = TRUE, data=flrcol)
anova(flrcol.cap, by="term")
```

## Spectrum vs. temp

```{r flrcol_avg}
aggplot(procspec(flrspec, opt="smooth", span=0.08), flrcol$temp, legend= T, main="Raw data")
aggplot(flrspec.nm.zero, flrcol$temp, legend= T,main= "Divided by max")
aggplot(flrspec.sum.zero, flrcol$temp, legend= T, main="Divided by sum")
```


# Phenology and fly eggs

```{r phenology_inventory}
phen.raw %>% count(date, plantid) %>% ggplot(aes(x=date, y=plantid))+ geom_tile() + geom_vline(xintercept=ymd("2021-06-28"))
```


## Snowmelt effects

```{r phenology_snow}
phen %>% drop_na(snow,temp) %>% group_by(snow, temp, julian) %>% summarize_at(c("open","buds"), mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=open, shape=snow, linetype=snow, color=temp))+ geom_point() + 
  geom_path(aes(group=paste(snow, temp))) +
  labs(y="Mean total flowers", 
       x="Day of year", color="Temperature", shape="Snowmelt", linetype="Snowmelt") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_color_brewer(palette="Set1", direction=-1) +
  scale_linetype_manual(values=c(2,1)) + theme_minimal() + 
  theme(legend.key.size = unit(2, "lines"))

phen %>% drop_na(snow,temp)%>% group_by(snow, temp, julian) %>% summarize_at("flowering", mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=flowering, shape=snow, linetype=snow, color=temp))+ geom_point() + 
  geom_path(aes(group=paste(snow, temp))) +
  labs(y="Proportion flowering", 
       x="Day of year", color="Temperature", shape="Snowmelt", linetype="Snowmelt") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_y_continuous(labels = scales::percent)+
  scale_color_brewer(palette="Set1", direction=-1) +
  scale_linetype_manual(values=c(2,1)) + theme_minimal() + 
  theme(legend.key.size = unit(2, "lines"))

phen %>% drop_na(snow,temp)%>% group_by(snow, temp, julian) %>% summarize_at("eggs", mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=eggs, shape=snow, linetype=snow, color=temp))+ geom_point() + 
  geom_path(aes(group=paste(snow, temp))) +
  labs(y="Mean total eggs", 
       x="Day of year", color="Temperature", shape="Snowmelt", linetype="Snowmelt") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_color_brewer(palette="Set1", direction=-1) +
  scale_linetype_manual(values=c(2,1)) + theme_minimal() + 
  theme(legend.key.size = unit(2, "lines"))

phen%>% drop_na(snow,temp) %>% group_by(year, snow, temp, julian) %>% summarize_at("has_egg", mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=has_egg, shape=snow, linetype=snow, color=temp))+ geom_point() + 
  geom_path(aes(group=paste(snow, temp))) +
  labs(y="Proportion of plants with eggs", 
       x="Day of year", color="Temperature", shape="Snowmelt", linetype="Snowmelt") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_y_continuous(labels = scales::percent)+
  scale_color_brewer(palette="Set1", direction=-1) +
  scale_linetype_manual(values=c(2,1)) + theme_minimal() + 
  theme(legend.key.size = unit(2, "lines"))
```

## Temperature effects

```{r phenology_temp}
ggplot(phen %>% drop_na(height_cm, temp), aes(x=julian, y=height_cm, color=temp))+ geom_point() + geom_path(aes(group=plantid))+
  labs(y="Plant height", x="Day of year", color="Temperature") +
  scale_x_continuous(breaks=seq(120,240, by=20))+ 
  scale_color_brewer(palette="Set1", direction=-1) +theme_minimal() 

phen %>% drop_na(snow,temp) %>% group_by(temp, julian) %>% drop_na(open, buds) %>% 
  summarize(openbuds.se = sd(open+buds)/sqrt(length(open+buds)), openbuds=mean(open+buds)) %>% 
  ggplot(aes(x=julian, y=openbuds, color=temp))+ geom_point() + 
  geom_path(aes(group=temp)) + geom_linerange(aes(ymin=openbuds-openbuds.se, ymax=openbuds+openbuds.se))+
  labs(y="Mean total open flowers + elongated buds", 
       x="Day of year", color="Temperature") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal()
lmer(I(open+buds) ~ factor(date) * temp + (1|plantid), data=phen.raw) %>% anova %>% kable(caption="mean open and buds")

phen %>% drop_na(temp)%>% group_by(temp, julian) %>% summarize_at("flowering", mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=flowering, color=temp))+ geom_point() + 
  geom_path(aes(group=temp)) + 
  labs(y="Proportion flowering", 
       x="Day of year", color="Temperature") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_y_continuous(labels = scales::percent)+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal()

phen %>% drop_na(temp) %>% group_by(temp, julian) %>% drop_na(eggs) %>% 
  summarize(eggs.se = sd(eggs)/sqrt(length(eggs)), eggs=mean(eggs)) %>% 
  ggplot(aes(x=julian, y=eggs, color=temp))+ geom_point() + 
  geom_path(aes(group=temp)) + geom_linerange(aes(ymin=eggs-eggs.se, ymax=eggs+eggs.se))+
  labs(y="Mean total eggs", 
       x="Day of year", color="Temperature") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal() 
lmer(eggs ~ factor(date) * temp + (1|plantid), data=phen.raw) %>% anova %>% kable(caption="mean eggs")

phen %>% drop_na(temp) %>% group_by(year, temp, julian) %>% summarize_at("has_egg", mean, na.rm=T) %>% 
  ggplot(aes(x=julian, y=has_egg, color=temp))+ geom_point() + 
  geom_path(aes(group=temp)) +
  labs(y="Proportion of plants with eggs", 
       x="Day of year", color="Temperature") +
  scale_x_continuous(breaks=seq(120,240, by=20))+
  scale_y_continuous(labels = scales::percent)+
  scale_color_brewer(palette="Set1", direction=-1) + theme_minimal() 
```

# Selection

##  Seed production

```{r selection, fig.width=7, fig.height=7}
all.mean %>% select(all_of(c(floraltraits, "seeds_est"))) %>% 
  pivot_longer(any_of(floraltraits), names_to="trait") %>% 
  ggplot(aes(x=value, y=seeds_est, color=temp)) + geom_point() + geom_smooth(method="lm",se=F, formula = y ~ poly(x, 2, raw=TRUE) ) + 
  facet_wrap(vars(trait), scales="free_x", labeller=as_labeller(traitnames.units)) + 
  labs(x="Trait", y=traitnames["seeds_est"], color="Temperature")+
  theme_bw() + scale_color_brewer(palette="Set1", direction=-1) + theme(legend.position = "top")

#get the the standardized selection differential S' (change in relative fitness / standardized trait value)
mt.mod.seeds <- map(set_names(floraltraits), 
                    ~ lm(seeds_rel ~ trait.z * temp, 
                         data=all.mean %>% as.data.frame() %>% mutate(trait.z = all.mean[[.x]] / traits.sd[[.x]]) %>% 
                           group_by(temp) %>% mutate(seeds_rel = seeds_est/mean(seeds_est, na.rm=T)))) # mean seeds in each temp
mt.mod.seeds.coefs <- map_dfr(mt.mod.seeds, tidy, .id="trait")
mt.mod.seeds.tests <- map(mt.mod.seeds, car::Anova, type=3) %>% map_dfr(tidy, .id="trait")
mt.mod.seeds.emm <- map_dfr(mt.mod.seeds, ~summary(emtrends(.x, ~temp, var="trait.z")), .id="trait")
mt.mod.seeds.tests %>% filter(!term %in% c("(Intercept)","Residuals")) %>% kable(caption="Seeds selection vs. temperature", digits=3)
ggplot(mt.mod.seeds.emm, aes(x=trait, color=temp, y=trait.z.trend, ymin=trait.z.trend-SE, ymax=trait.z.trend+SE)) + 
  geom_hline(yintercept=0)+ geom_pointrange(position=position_dodge(width=0.3)) + 
   theme_bw() + scale_color_brewer(palette="Set1", direction=-1) + theme(legend.position = "top") + 
  labs(x="", y="Standardized selection differential, total seeds", color="Temperature") +
  scale_x_discrete(labels=traitnames)
```


## Fly damage

```{r selection_fly , fig.width=7, fig.height=7}
all.mean %>% select(all_of(c(floraltraits, "prop_infested"))) %>% 
  pivot_longer(any_of(floraltraits), names_to="trait") %>% 
  ggplot(aes(x=value, y=prop_infested, color=temp)) + geom_point() + geom_smooth(method="glm",se=F, method.args = list(family="binomial")) + 
  facet_wrap(vars(trait), scales="free_x", labeller=as_labeller(traitnames.units)) + 
  labs(x="Trait", y=traitnames["prop_infested"], color="Temperature")+
  theme_bw() + scale_color_brewer(palette="Set1", direction=-1) + theme(legend.position = "top")

mt.mod.infested <- map(set_names(floraltraits), 
                    ~ lm(infested_rel ~ trait.z * temp, 
                         data=all.mean %>% as.data.frame() %>% mutate(trait.z = all.mean[[.x]] / traits.sd[[.x]]) %>% 
                           group_by(temp) %>% mutate(infested_rel = prop_infested/mean(prop_infested, na.rm=T)))) # mean infested in each temp
mt.mod.infested.coefs <- map_dfr(mt.mod.infested, tidy, .id="trait")
mt.mod.infested.tests <- map(mt.mod.infested, car::Anova, type=3) %>% map_dfr(tidy, .id="trait")
mt.mod.infested.emm <- map_dfr(mt.mod.infested, ~summary(emtrends(.x, ~temp, var="trait.z")), .id="trait")
mt.mod.infested.tests %>% filter(!term %in% c("(Intercept)","Residuals")) %>% kable(caption="Fly selection vs. temperature", digits=3)
ggplot(mt.mod.infested.emm, aes(x=trait, color=temp, y=trait.z.trend, ymin=trait.z.trend-SE, ymax=trait.z.trend+SE)) + 
  geom_hline(yintercept=0)+ geom_pointrange(position=position_dodge(width=0.3)) + 
   theme_bw() + scale_color_brewer(palette="Set1", direction=-1) + theme(legend.position = "top") + 
  labs(x="", y="Standardized selection differential, proportion infested", color="Temperature") +
  scale_x_discrete(labels=traitnames)
```

